Marginally-Stable Thermal Equilibria of Rayleigh-Bénard Convection

### 2021-04-07 publication_materials/ms_rbc.tex
intro rev2a

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in large-scale astrophysical and geophysical settings. 
Our understanding of turbulent convective systems is generally limited by the prohibitive computational costs associated with direct numerical simulations \cite{Zhu}. 
There exists a substantial body of work concerning the asymptotic scaling laws of such systems with no general consensus. 
Reduced models are particularly useful in this context because they allow us to study the underlying behavior of turbulent convection with relatively low expense.

@@

Reduced models are often valuable when supplemented by direct simulation [source?]. 
Nonlinear simulations performed with quasilinear solutions as initial conditions do not undergo a convective transient period of turbulent upwelling. 
Instead, the characteristic jet-like structure is somewhat present on initialization. 
This is practical because the convective transient period generally involves high-Reynolds number flows whose resolution is prohibitive.

@@

The most resilient and unexpected feature of thermally equilibrated mean temperature profiles are the pronounced dips adjacent to the boundary layers. 
These dips appear in every solution, regardless of $Ra$. 
Physically, they correspond to thin layers in which the temperature gradient reverses, contradicting a hypothesis of \cite{Malkus}. 
This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure \ref{fig:flux}. 
\begin{figure}
    \centering
    \includegraphics[width=3.4in]{T_profs_na.png}
    \caption{mean temperature profiles $\bar{T}$ for $Ra = 10^8$. 
    The initial profile is given by (\ref{EQ:T0}). 
    The equilibrated curve refers to the mean temperature profile of the marginally-stable thermally-equilibrated state. 
    The simulation profile is obtained by performing a 2-dimensional nonlinear simulation with \texttt{Dedalus} until thermal and kinetic relaxation. 
    The simulation temperature data are horizontally- and time-averaged. 
    The initial, equilibrated, and simulation profiles have increasingly relaxed boundary regions (respectively). 
    The equilibrated profile exhibits prominent dips, nested alongside the boundary regions. 
    The purpose of this feature is not well understood.}
    \label{fig:T0_profiles}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=3.4in]{kx_m_ra1.png}
    \caption{Wavenumbers of marginally-stable modes of thermally equilibrated states. 
    Markers are color-coded according to their adjacent local maxima index in the eigenvalue spectrum. 
    For example, the spectrum corresponding to $Ra = 10^5$ has marginal wavenumbers $k_x = \pi, \, 1.5\pi$ adjacent to a presumed local maximum between these two allowed values. 
    The $Ra = 10^9$ spectrum, shown in lower right corner of Figure \ref{fig:flux}, has three presumed local maxima, with a single marginal mode adjacent to the first maximum ($k_x = 1.5\pi$), two marginal modes adjacent to the second maximum ($k_x = 6\pi, \, 6.5\pi$), and two marginal modes adjacent to the third maximum ($k_x = 23.5\pi, \, 24\pi$). 
    The largest wavenumbers of the green branch obey a power-law relationship with $Ra$}
    \label{fig:kx_marginals}
\end{figure}
\begin{figure*}
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.png}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.png}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states $Ra = 2 \times 10^5$ (top) and $Ra = 10^9$ (bottom). 
    Advection profiles belong to marginally-stable modes. 
    For low $Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. 
    For large $Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. 
    A combination of of progressively wider advection profiles is necessary to transition to the $k_x = 1.5\pi$ mode.}
    \label{fig:flux}
    \vspace{0.4in}
    \includegraphics[width=7.1in]{nu_ra.PNG}
    \caption{Nusselt numbers are shown for marginally-stable thermally-equilibrated states, statistically-steady direct numerical simulations performed with \texttt{Dedalus} \cite{Anders_cd}, and equilibrated nonlinear solutions, which maximize heat flux, given by \cite{Waleffe}. 
    Both obey power-law relationships with $Ra$, with $Ra \sim \Nu^3$ in agreement with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} as well as \cite{Malkus}. 
    Marginally-stable states' Nusselt numbers exceed simulation Nusselt numbers. 
    This can be well explained by the contrasting boundary region geometries shown in Figure \ref{fig:T0_profiles}.}%
    \label{fig:nu_vs_ra}%
\end{figure*}

@@

\bibitem{Yalniz} Coarse graining the state space of a turbulent flow using periodic orbits Yalnız, Gökhan; Hof, Björn; Burak Budanur, Nazmi

>>>>

\section{Introduction}
Rayleigh-B\'enard convection plays a foundational role in large-scale astrophysical and geophysical settings.
Such systems regulate heat transfer and can be associated with violently turbulent flows.

Aggressive convection is associated with large Rayleigh numbers $Ra$ and is difficult to simulate. State of the art convection simulations performed by \cite{Zhu} have reached $Ra = 10^{14}$ but estimates for stars are $Ra \sim 10^{20}$ \cite{Ossendrijver}. 
The scaling behavior of the Nusselt number $\Nu$ (convective strength) in the highly turbulent ultimate regime is of particular interest.
There exists a substantial body of work pertaining to this specific topic with no general consensus \cite{Malkus}, \cite{Kraichnan} \cite{Spiegel}, \cite{Castaing}, \cite{Grossman}, \cite{Ahlers}.

Reduced models are particularly useful in this context because they allow us to study the system with relatively low expense. 
Much of the published work on this topic relates to unstable time-invariant solutions \cite{Waleffe}, \cite{Sondak}, \cite{Wen}, \cite{chini_cells}. 
Simulations and analysis performed by \cite{Yalniz} suggest that solution trajectories can be described by a Markov process, vacillating between these steady solutions.

@@

@@

The most resilient and unexpected feature of thermally equilibrated mean temperature profiles are the pronounced dips adjacent to the boundary layers. 
These dips appear in every solution, regardless of $Ra$. 
Physically, they correspond to thin layers in which the temperature gradient reverses, contradicting a hypothesis of \cite{Malkus}. 
This counter-diffusion, which opposes overall heat transfer, is overcome by the coinciding advective flux, shown in Figure \ref{fig:flux}. 
\begin{figure}
    \centering
    \includegraphics[width=3.4in]{T_profs_na.png}
    \caption{mean temperature profiles $\bar{T}$ for $Ra = 10^8$. 
    The initial profile is given by (\ref{EQ:T0}). 
    The equilibrated curve refers to the mean temperature profile of the marginally-stable thermally-equilibrated state. 
    The simulation profile is obtained by performing a 2-dimensional nonlinear simulation with \texttt{Dedalus} until thermal and kinetic relaxation. 
    The simulation temperature data are horizontally- and time-averaged. 
    The initial, equilibrated, and simulation profiles have increasingly relaxed boundary regions (respectively). 
    The equilibrated profile exhibits prominent dips, nested alongside the boundary regions. 
    The purpose of this feature is not well understood.}
    \label{fig:T0_profiles}
\end{figure}
\begin{figure}
    \centering
    \includegraphics[width=3.4in]{kx_m_ra1.png}
    \caption{Wavenumbers of marginally-stable modes of thermally equilibrated states. 
    Markers are color-coded according to their adjacent local maxima index in the eigenvalue spectrum. 
    For example, the spectrum corresponding to $Ra = 10^5$ has marginal wavenumbers $k_x = \pi, \, 1.5\pi$ adjacent to a presumed local maximum between these two allowed values. 
    The $Ra = 10^9$ spectrum, shown in lower right corner of Figure \ref{fig:flux}, has three presumed local maxima, with a single marginal mode adjacent to the first maximum ($k_x = 1.5\pi$), two marginal modes adjacent to the second maximum ($k_x = 6\pi, \, 6.5\pi$), and two marginal modes adjacent to the third maximum ($k_x = 23.5\pi, \, 24\pi$). 
    The largest wavenumbers of the green branch obey a power-law relationship with $Ra$}
    \label{fig:kx_marginals}
\end{figure}
\begin{figure*}
    \centering
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{flux_sup_n.png}
    \end{tabular}
    \begin{tabular}{@{}c@{}}
        \includegraphics[width=3.4in]{EV_spectra_2ra.png}
    \end{tabular}
    \caption{Heat fluxes (left) and eigenvalue spectra (right) of equilibrated states $Ra = 2 \times 10^5$ (top) and $Ra = 10^9$ (bottom). 
    Advection profiles belong to marginally-stable modes. 
    For low $Ra$, a single mode with $k_x = 1.5\pi$ is sufficient to oppose boundary layer diffusion and facilitate heat flux throughout the bulk of the domain. 
    For large $Ra$, high-wavenumber modes contribute pronounced small-scale advection profiles which tightly hug the thin boundary layers. 
    A combination of of progressively wider advection profiles is necessary to transition to the $k_x = 1.5\pi$ mode.}
    \label{fig:flux}
\end{figure*}
\begin{figure*}
    \centering
    \includegraphics[width=7.1in]{nu_ra.PNG}
    \caption{Nusselt numbers are shown for marginally-stable thermally-equilibrated states, statistically-steady direct numerical simulations performed with \texttt{Dedalus} \cite{Anders_cd}, and equilibrated nonlinear solutions, which maximize heat flux, given by \cite{Waleffe}. 
    Both obey power-law relationships with $Ra$, with $Ra \sim \Nu^3$ in agreement with Figures \ref{fig:bl_ra}, \ref{fig:kx_marginals}, and \ref{fig:del_inv} as well as \cite{Malkus}. 
    Marginally-stable states' Nusselt numbers exceed simulation Nusselt numbers. 
    This can be well explained by the contrasting boundary region geometries shown in Figure \ref{fig:T0_profiles}.}%
    \label{fig:nu_vs_ra}%
\end{figure*}

@@

\bibitem{Yalniz} Yalnız, Gökhan; Hof, Björn; Burak Budanur, Nazmi. Coarse graining the state space of a turbulent flow using periodic orbits

\bibitem{Ossendrijver} The solar dynamo. Ossendrijver, M. The Astron. Astrophys. Rev. 11, 287–367 (2003).

\bibitem{Kraichnan} Kraichnan, R. Turbulent Thermal Convection at Arbitrary Prandtl Number (1962). https://doi.org/10.1063/1.1706533

% \bibitem{Featherstone} The Spectral Amplitude of Stellar Convection and its Scaling in the High-Rayleigh-Number Regime. Featherstone, N. A., \& Hindman, B. W. (2016). The Astrophysical Journal, 818, 32–45.

\bibitem{chini_cells} Large Rayleigh number thermal convection: heat flux predictions and strongly nonlinear solutions. Chini, G. P., \& Cox, S. M. (2009). Physics of Fluids, 21, doi:10.1063/1.3210777

\bibitem{Sondak} Sondak, D., Smith, L., \& Waleffe, F. (2015). Optimal heat transport solutions for Rayleigh–Bénard convection. Journal of Fluid Mechanics, 784, 565-595. doi:10.1017/jfm.2015.615

% scaling review
\bibitem{Grossman} Grossman, S., \& Lohse, D. (2000). Scaling in thermal convection: A unifying theory. Journal of Fluid Mechanics, 407, 27-56. doi:10.1017/S0022112099007545

% modal convection: scaling
\bibitem{Spiegel} Thermal Turbulence at Very Small Prandtl Number. Spiegel, E. (1962) Journal of Geophysical Research, https://doi.org/10.1029/JZ067i008p03063

\bibitem{Castaing} Castaing, B. et al. (1988) Scaling of hard Thermal Turbulence in Rayleigh-B\'enard Convection

\bibitem{Ahlers} Ahlers, G. Heat transfer \& large-scale dynamics in turbulent Rayleigh-B´enard convection

\bibitem{Shraiman} Shraiman BI, Siggia ED. Heat transport in high-Rayleigh-number convection. Phys Rev A. 1990 Sep 15;42(6):3650-3653. doi: 10.1103/physreva.42.3650. PMID: 9904456.

\bibitem{Wen} Wen, B., Goluskin, D., Doering, C. (2020) Steady Rayleigh–B´enard convection between no-slip boundaries
